# (PART\*) COMMUNITY ANALYSIS AND FEEDBACK {-}

```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

# Scientific Posters

In this section we will go over how to create a scientific poster. We chose a poster as our final deliverable in C-MOOR because they:


- **Facilitate participation in research symposiums**: Students can use their poster to apply to present at research symposiums or conferences with minimal additional effort beyond what they do in class. If your institution does not have a research day, consider starting one using these posters as a foundation as we did at Clovis Community College! Your school's library is another great point of contact in getting a poster session or research day set up. Some of the events students have participated in using their C-MOOR projects include:

  - The Clovis Community College Research Day
  - The Notre Dame of Maryland University Research Day
  - The GRADS-4C Conference (2025)

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Image collage of 3 photographs showing students in front of their research posters at various symposia"}

ottrpal::include_slide("https://docs.google.com/presentation/d/1dI8-_iVqbkzNMf11M4dK85E8ZW3OyZECs_YwMKw5fhs/edit#slide=id.g362974128df_0_0")

```

- **Serve as a physical representation of student achievement**: If printed, once finished with the class, students can choose to take their poster with them or leave it with the class to be hung up around the classroom or the hallways. This gives their research project more visibility, serves as an advertisement for other students who might be interested in taking the class, and showcases what students have accomplished. 


- **Can easily be shared digitally**: A presentation can't be repeated twice without a recording, and each recording can take up a lot of digital storage. A poster on the other hand, is a great compressed product that can be posted on a student's (or the institution's) social media and help get their work seen by others. Posters are even small enough that they can be attached to an email and sent to prospective supervisors and collaborators interested in a quick summary of their work. See our [Look at This!](https://help.c-moor.org/c/look-at-this/) category on the C-MOOR Academy Discussion Forum to see all the posters students have created and shared so far!

- **Posters can be used to measure student achievement and understanding**: We have previously used these posters to evaluate student learning, even years after students have left the classroom. Check out our [poster on posters!](https://drive.google.com/file/d/1kK6FBBLbHiAAsTUgxwzHRlK1qCo__eCy/view?usp=sharing)

## Making a Group Poster 

### Preparing to make a poster

An important part of scientific research is presenting your findings. 
Poster is a powerful visual way to communicate new and exciting findings, share ideas and get feedback. 
No wonder Poster sessions are an integral part of any conference or symposium. 
Throughout research project work, students will work with their groups to put together a scientific research poster.
 

### Part 1 - Choose a Template
*Estimated time: 5 minutes*

1. With your group, open a suggested [poster template](https://docs.google.com/presentation/d/1chz02nzYklAEjOtHrPsJZccGy1sFUZvqdlfcUyfj3I4/edit?usp=sharing) to use for your poster. 
1. Discuss with your group how you will divide up the work and exchange important information (e.g. phone numbers, email). Consider the following sections: 

- Abstract/Introduction
- Methods
- Results
- Conclusions/Discussion
- References 
- Acknowledgments

Notice these posters are sized at 36 x 42 inches, but there are many other commonly used sizes. 
Refer to your instructor on what poster size to use; not every poster will fit every stand when printed (if applicable). 
Notice that the posters in this slidedeck come in two-column, three-column, and mixed formats.
Think about your figures and tables and select a format that makes the most sense for them. 
There are also some example posters in the last two slides that we will use to examine the anatomy of a research poster further in the next part.


### Part 2 - Make an Academic Research Poster

*Estimated time: ---*

Complete the following components of your research poster with your group. You might not do them in this order, but these are the components you are being graded on. For more details on each of these sections and their role in a scientific paper, see the Scientific Literature Lab.

#### The header: title, authors, and affiliations

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Image showing 4 examples of headers from C-MOOR associated posters. Each header has the title of the poster in the largest font, the logos of supporting organizations, and the author list. The headers differ in color and overall aesthetic style."}

ottrpal::include_slide("https://docs.google.com/presentation/d/1hYKF7Ss3vJ8rrUIH7ByNh1BUlRa2fhsJhq8MXzEowCc/edit?slide=id.g33d6e3b928f_0_67#slide=id.g33d6e3b928f_0_67")

```

Since you completed this work as a team of scientists, all of your team members are considered authors. List your team members in alphabetical order. The author’s affiliation is the university, college, research institution or company that the work was conducted at.  

- An example affiliation for Clovis Community College would be: “Department of Biology, Clovis Community College, California, United States.” 
- If multiple institutions exist, you will need to use superscript to denote who has what affiliation (see the example poster).

Include the C-MOOR logo, the logo of your institution(s), and the most immediately relevant funding source (if applicable) in the header.

<!-- NOTE: &nbsp; creates a space in Markdown. In this instance, adding extra spaces after the checkbox didn't seem to work. -->

**Check the header of your poster:**

```
- All authors are listed
- All authors’ home institutions are referenced
- The title of the poster is short and focused on your specific project topic
- Appropriate logos are included: E.g. the C-MOOR logo, the logo of your institution
- Relevant funding source (if applicable) is listed
```

A note on the title:

- The title of the poster describes the main result of your research
- The title of the poster does not overstate the findings or significance of your research


#### Abstract

C-MOOR Abstract examples:

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Image showing 2 abstracts from C-MOOR associated posters. One has a header that reads Abstract."}

ottrpal::include_slide("https://docs.google.com/presentation/d/1dI8-_iVqbkzNMf11M4dK85E8ZW3OyZECs_YwMKw5fhs/edit#slide=id.g36d8202a6d3_0_0")

```

An abstract is a concise summary of your work. An effective abstract will be well designed and communicated, and will inform the reader of the relevant background to the research, scientific hypothesis being tested, the purpose of the study, the main methods, and the most important results and conclusions.  

Abstract can be difficult to write because it combines all research pieces and requires effective and consice communication of those sections.
Many scientists choose to write the abstract last, after they fully understand the conclusions and implications of their work. 

**Your abstract should do the following:**


1. **Clearly and briefly** state purpose, hypothesis, aims, and/or objectives of the research.

```
- e.g. Our hypothesis is that the healthy gut microbiome has more microbial diversity than the gut of an individual with a celiac disease.`
- e.g. We hypothesize that maternal antibiotic treatment correlates with higher antibicrobial resistance in infant microbiome.`
```

2. **Clearly and briefly** summarize relevant background information such as: 

```
- A problem or question at hand 
- Model organism
- The gene of interest and its function 
- The relevant phenotypes (e.g. disease-state) 
- The known connection between genotype and phenotype
- Relevant biological processes
```

3. **Clearly and briefly** summarize the main results of your study, and how they may relate to the hypothesis.

```
- e.g. We found that gene X was differentially abundant between the eye and all other tissues, suggesting gene X plays an important role in the eye. This may be important for learning more about eye development and X condition.` 
- e.g. We found an association between age and the presence of Y bacteria, which supports our hypothesis that Y bacteria is involved in the disease state`
```

4. **Clearly and briefly** summarize methodology you used in your research.

```
- e.g. Using Galaxy we analyzed the genomic diversity of the gut microbiome and compared fecal samples between individuals with Celiac Disease (CD) and a control group on a healthy diet without CD.`
- e.g. We analyzed a publically available dataset comparing RNA-seq gene expression between the left and right eye in DESeq2.`
```

***Overall, the Abstract is heavy on summary aspect of things, and lean on details aspect of things***.

#### Introduction

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Image showing 2 introductions from past C-MOOR posters. Both contain a header text that reads: Introduction, and both contain figures, one of which contains a figure legend."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1dI8-_iVqbkzNMf11M4dK85E8ZW3OyZECs_YwMKw5fhs/edit#slide=id.g362974128df_0_18")
```

An introduction typically appears immediately after the Abstract section, and while abstract is very brief and consise, Introduction provides more context and background information to your research. 
Indeed, the main goal of the Introduction is to **clearly elaborate on relevant background information and provide context for research gap and hypothesis**.

***Overall, the Introduction is longer than the Abstract, heavier on providing research context and background, and leaner on result summary and conclusions.***.

- Sometimes, you will se an Introduction section used instead of the Abstract + Introduction sections, in which case, the Introduction will contain both, abstract information plus additional background information, such as this published PacBio's  abstract on 'Genome-resolved metagenome assembly of human oral microbiome using highly accurate long-read sequencing'. 

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Example Introduction from the PacBio Website, including a header that reads Introduction."}

ottrpal::include_slide("https://docs.google.com/presentation/d/1hYKF7Ss3vJ8rrUIH7ByNh1BUlRa2fhsJhq8MXzEowCc/edit?slide=id.g33d6e3b928f_0_149#slide=id.g33d6e3b928f_0_149")

```

#### Materials and Methods

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Image showing two introductions from previous C-MOOR posters. Each has a header titled Methods and contain figures and tables, with some legends included. The example on the right also contains text in paragraph form."}

ottrpal::include_slide("https://docs.google.com/presentation/d/1hYKF7Ss3vJ8rrUIH7ByNh1BUlRa2fhsJhq8MXzEowCc/edit?slide=id.g33d6e3b928f_1_1#slide=id.g33d6e3b928f_1_1")

```

The materials and methods section will detail your analysis of the data. Don’t provide any of your results, just the methods. If you did not generate the data yourself from raw samples, you will simply cite the paper that made them instead of detailing their construction. We will also need to list any programs we used and provide credit to their creators.

Some other things you might include would be what type of analysis you decided to do (which parts of the body parts you analyzed, sets of genes, what p-value you used, etc.).

#### Results

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "An image showing an example of a short results section. The header describes the main finding of the figure, with smaller body text providing more explanation. The figure is multi-panel and has a legend below it which has a bolded title and a description for each panel."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1hYKF7Ss3vJ8rrUIH7ByNh1BUlRa2fhsJhq8MXzEowCc/edit?slide=id.g33d6e3b928f_0_6#slide=id.g33d6e3b928f_0_6")
```

The Results section is where you will detail your data primarily through **figures and tables**, though sometimes written text is included. Begin by creating your tables and figures. 

**Figure body**:

- Has high resolution images, well spaced and labeled parts and text. 
- Place the figures and tables in order of how you want to present them and name them such as **Figure 1, Figure 2, Table 1, Table 2, etc**.

**Figure Legend**:

- Figures have their legends *underneath* them. Tables may have their legends also *above* them.
- The legend should be in a smaller font than the main text on the poster.
- A legend should includes a **figure title**, which is a declarative statement that summarizes  findings.
- A legend also includes **figure text** which a) identifies (and can  briefly describe  parts), b) includes brief description of of methods necessary to understand figure and c) should include relevant statistics (stats).

**Optional components**

- Figures and Tables can have additional (optional) text. E.g.an optional text can describe the findings of the figure upfront to engage.
- A **Section Title/Header** which summarizes section or figure in a manner that is broader than figure legend title - an attention grabber.
- Bullet pointed text that summarizes the main findings of the figures/tables.

#### Conclusions/Discussion

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "An image showing 4 example conclusion or discussion sections from prior C-MOOR posters. Some have headers, but the headers differ in style. One header is more descriptive and others just say Discussion. All 4 have text in bullet point form, some with references in them."}

ottrpal::include_slide("https://docs.google.com/presentation/d/1dI8-_iVqbkzNMf11M4dK85E8ZW3OyZECs_YwMKw5fhs/edit#slide=id.g362974128df_0_52")

```

The discussion section of the paper is your chance to analyze and interpret your results. The discussion section generally addresses the following:

- What do your results mean?
- How do they fit into the bigger picture?
- If any experiments did not give expected results, hypothesize why that might have been the case and propose alternate experiments that could confirm or clarify your results.
- Include at least one sentence of future work that you would do if you had more time or what students in upcoming semesters could do to continue to answer your questions.

#### References

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "An example reference section from a prior C-MOOR poster, with a bolded header that reads References followed by a list of references in a consistent style."}

ottrpal::include_slide("https://docs.google.com/presentation/d/1dI8-_iVqbkzNMf11M4dK85E8ZW3OyZECs_YwMKw5fhs/edit#slide=id.g362974128df_0_80")

```

All the references that you cite on your poster must be present in a References section including the following sections: Introduction, Methods, and Discussion. To save space on our posters, we will number our references (ex. 1-5) and use the numbers as citations throughout the text of your poster. You may have a lot of references; it is okay to put them in tiny text if you have to in order to make them fit.

There are many different ways to format the reference section. We will make ours in alphabetical order by the first author’s last name. All of your sources must be scientific journals and should use the following format:

> Authors (year) “Title.” Journal Name, vol. #, page #s, DOI

Online article that is also in print:

> Haussecker D., Huang Y., Lau A., Parameswaran P., Fire A. Z. and M. A. Kay (2010) “Human tRNA-derived small RNAs in the global regulation of RNA silencing.” RNA, Vol. 16, page 637-695,  doi:10.1261/rna.2000810

Online article only:

> Marianes, A. and A. C. Spradling (2013) “Physiological and stem cell compartmentalization within the Drosophila midgut.” eLife, doi:10.7554/eLife.00886


#### Acknowledgements

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "An image of 3 different acknowledgement sections from prior C-MOOR posters. All have a header that reads: Acknowledgements. One contains a grant number and icons from various supporting institutions."}

ottrpal::include_slide("https://docs.google.com/presentation/d/1dI8-_iVqbkzNMf11M4dK85E8ZW3OyZECs_YwMKw5fhs/edit#slide=id.g362974128df_0_30")

```

The acknowledgements section is where you give thanks to the people, organizations, and institutions that have supported you in your research. If relevant, include the grant ID # of your funding source. Institutions and organizations that have contributed to your research - but you do not belong to - can be thanked here.

### Part 3 - Proofread and Add Final Touches

*Estimated time: 30 min to an hour*

1. Each group member should re-read the poster from beginning to end and fix any typos or grammatical errors.
1. Check the alignment of figures, text boxes, titles, etc.
1. Add some finishing touches. You can play with the color, the font, add additional images if it’s relevant.

### Part 4 - Canvas Discussion

*Estimated time: 30 min*

You will turn in your poster to be graded as a group in a Canvas Assignment and post it to a Canvas Discussion to be viewed by the class.

1. Convert your poster to a pdf.
1. Have one member of your group turn in the pdf of your poster to the Graded Canvas Assignment. This assignment is already set up so that if one group member turns it in, it will show as submitted for all students in the group. This is where your instructor will grade you poster as a group.
1. Have one member of your group post a pdf of your poster in the Canvas C-MOOR Poster Discussion.
    a. With your poster, introduce your  group members and copy and paste your abstract into the post.
    a. Insert your pdf into the post and edit the link so that it automatically shows the inline preview. This will make it easier for students to view your poster.
1. As an individual, read through the other posters from different groups.
1. Post comments

#### Grading Criteria

Your instructor will provide for you a rubric specific to your class. 
Refer to the rubric as your build your poster to make sure you have all the neccessary components.

#### Footnotes

**Resources**

- [Google Doc]
- [Generic rubric](https://docs.google.com/document/d/13eB1fwHEB1I00JHBUQLZIJsi-iQZYaWaXThPqCTU17g/edit?usp=sharing)

**References**

- [Abstract vs. Introduction](https://fourwaves.com/blog/abstract-vs-introduction/)
- [Preparing and presenting effective research posters] (https://onlinelibrary.wiley.com/doi/10.1111/j.1475-6773.2006.00588.x)

**Contributions and Affiliations**

- Stephanie R. Coffman, Clovis Community College
- Valeriya Gaysinskaya, Johns Hopkins University
- Frederick Tan, Johns Hopkins University
- Sayumi York, Notre Dame of Maryland University

Last Revised: August 2025
