
# (PART\*) Project: 16S rRNA Research {-}

<!-- Set up code of OTTR Book-->

```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```



# Reference - Meet the MISO study

Let's get prepared to do some research using 16S rRNA data! In this section, we'll be exploring data from *Impact of a 7-day homogeneous diet on interpersonal variation in human gut microbiomes and metabolomes* by Guthrie et al. (2022). This study is one of many that explores the relationship between diet and the human gut microbiome and will help familiarize us with the format that 16S rRNA data takes. We'll start slow and look at just a few lines of data in Google Sheets and then use phyloseq and DESeq2 in R to take our analysis to the next level.

Human diet has been implicated heavily in the establishment and maintenance of the gut microbiome. For example, human babies undergo a drastic change in the gut microbiome following the transition to solid food.

<!-- REFs: Xu et al. (2015) -->

It's difficult to capture the mechanisms and effects of diet on the gut microbiome given the sheer number of variables involved. The microbiome individuality and stability over time (MISO) study aimed to explore the connection between diet, microbiome, and metabolites by looking at the effect of feeding a standardized diet to people over 7 days. 

### MISO study diet and sampling schedule

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "An image showing the sampling scheme of the MISO study. Samples are taken at 5 time points, 2 during the baseline diet phase, 2 during the homogenized diet phase, and one during the washout diet phase. The experiment lasts 28 days total. Some samplings are closer to each other than others temporally."}

ottrpal::include_slide("https://docs.google.com/presentation/d/1dEZtb0TgLbzhbr3HxQElwg8hENur2HMIzseg19OGd0g/edit#slide=id.g33d6fd7d4b6_0_27")

```

The figure above shows the study design for the MISO study. You do not need to memorize all these details; feel free to refer to this page throughout the project:

- Participants eat their usual, **baseline diet (BD) for 14 days**
- Participants all eat the same diet, the **homogenized diet (HD), for 7 days**
- Participants return to their usual diet during the **washout (WO) period for 7 days**

The study lasts a total of 28 days. Samples from the blood, stool, and urine, our metabolite and 16S rRNA data are taken at 5 different timepoints:

- **Timepoint 1:** Day 0 (the start of the study)
- **Timepoint 2:** Day 13
- **Timepoint 3:** Day 17
- **Timepoint 4:** Day 21
- **Timepoint 5:** Day 28


### MISO study variables and factors

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "A slide titled MISO study variables. There are 3 columns from left to right: study design represented by a calendar and clock, subject, represented by three people, and metabolites, represented by a partially filled sample tube rack. Under each section are related variables listed in the following study design variables section."}

ottrpal::include_slide("https://docs.google.com/presentation/d/1dEZtb0TgLbzhbr3HxQElwg8hENur2HMIzseg19OGd0g/edit#slide=id.g33d6fd7d4b6_0_61")

```

We have a number of variables we can use in our analysis. Notice that the variables for the study and subjects are in lowercase. This is also how you will access these variables in R.

#### Study design variables

<!-- How to align table using this format? -->

|Variable|What is it?|Factors|
|:----|--------------|----:|
|timepoint|The 5 samplings that occur on days 0, 13, 17, 21, and 28 coded as timepoints 1 through 5|1, 2, 3, 4, 5| 
|diet|The diet the subject was on during the sampling|BD, HD, WO|

#### Subject variables

|Variable|What is it?|Factors|
|:-|----|----:|
|subject|A unique ID given to each subject (participant)|S## (ex. S02 is subject 2. Note that while there are a total of 21 subjects in the study, but their subject numbers are not 1 through 20| 
|sample|A unique ID given to every sample taken during the study that includes the subject and timepoint of the sample|MISO-Subject##-Sample# (ex. MISO1-S02-1 is the sample from subject 2 at timepoint 1)|
|gender|The gender of the subject|M, F. All subjects were cisgender| 
|age|The age in years of the subject|A continuous variable from 23 to 75 years old| 
|bsa|Body surface area; a measure of body size|A continuous variable from 1.6 to 2.8| 

#### Metabolite variables

|Variable|What is it?|Factors|
|:----|--------------|----:|
|Creatinine|Creatinine|A continuous variable from 1072 to 3971| 
|PCS|*p*-cresol sulfate|A continuous variable from 2 to 95|
|IS|Indoxyl sulfate|A continuous variable from 3 to 58| 
|HIPP|Hippuric acid|A continuous variable from 16 to 1119| 
|PAG|Phenylacetylglutamine|A continuous variable from 16 to 318| 

### Amplicon sequence variants (ASVs) data

Finally we have our microbe count data, or our *Amplicon Sequence Variants (ASVs)*. Each ASV is a unique sequences that differs by as little as a single nucleotide from other ASVs and represents a specific microbe. We have assigned taxonomy to these microbes, but not all microbes have taxonomic data through the species level. These missing fields will appear as NA in the data.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "A slide titled Amplicon Sequence Variants (ASVs). On the side of the slide reads in bullet points: 1) ASVs are unique sequences that differ by as little as a single nucleotide and represent distinct microbes. 2) Taxonomic information has been assigned to each ASV at as high a resolution as possible. On the right side there is a bolded title that reads: ASVs are assigned taxonomic information from databases of known sequences and microbes, underneath which is an image of a laptop and three colored lines representing unique ASVs (sequences) that correlate to one of three microbes in correlated colors."}

ottrpal::include_slide("https://docs.google.com/presentation/d/1dEZtb0TgLbzhbr3HxQElwg8hENur2HMIzseg19OGd0g/edit#slide=id.g33d6fd7d4b6_0_136")

```

You may see the word Operational Taxonomic Unit (OTU) in phyloseq and in other published studies. **ASVs and OTUs have some differences between them but they both represent distinct units of microbial taxa**, although they have some key differences. 

Each sample is associated with a count of each ASV. We can compare the counts of these ASVs between samples to determine differences in the composition of the microbiome between samples.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "A slide titled Each sample has a count of each ASV. On the left side of the slide is an upside down pyramid with the largest rectangle on top and smallest on the bottom. From the top to bottom are the taxonomic ranks: Kingdom, Phylum, Class, Order, Family, Genus, and Species, along with the taxonomy for Roseburia inulinivorans. An arrow denotes that this ASV is ASV 53 and points to a table that relates the counts of each ASV found in each specific sample. For ASV 53, there are 32 counts found in the sample MISO1-S02-1 and 16 counts found in the sample MISO1-S23-1."}

ottrpal::include_slide("https://docs.google.com/presentation/d/1dEZtb0TgLbzhbr3HxQElwg8hENur2HMIzseg19OGd0g/edit#slide=id.g33d6fd7d4b6_0_113")

```


### Footnotes

#### Resources

#### Contributions and affiliations

- Sayumi York, Notre Dame of Maryland University


